---
title: "DATA 605 Homework2"
author: "Chinedu Onyeka"
date: "2023-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Questions: ** 
<center>
<h3> 1: Problem Set 1 </h3>
</center>
### 1.1
Show that $A^{T}A \neq AA^{T}$ in general. (Proof and demonstration).  
</b>  
**Solution**  

For the expression $A^{T}A = AA^{T}$ to be true in general it must be true for all A where A is a matrix. So, if we can get at least one A that does not satisfy the expression, then the expression is not true in general for all A.
</b>
Suppose a matrix $A$ is an $m \times n$ matrix, its transpose $A^{T}$ will be an $n \times m$ matrix

Let $A$ be a $2 \times 2$ square matrix:
$$
A = 
\begin{bmatrix}
    u & v \\
    x & y \\
\end{bmatrix}
$$
The transpose of A will be
$$
A^T = 
\begin{bmatrix}
    u & x \\
    v & y \\
\end{bmatrix}
$$

The multiplication of $A^T$ by $A$, $A^TA$ is given by
$$
A^TA = 
\begin{bmatrix}
    u^2+x^2 & uv+xy \\
    uv+xy & v^2+y^2 \\
\end{bmatrix}
$$
The multiplication of $A$ by $A^T$, $AA^T$ is given by

$$
AA^T = 
\begin{bmatrix}
    u^2+v^2 & ux+vy \\
    ux+vy & x^2+y^2 \\
\end{bmatrix}
$$
We can see that unless x = v, the two would not be equal.  
Therefore,

$$
\begin{bmatrix}
    u^2+x^2 & uv+xy \\
    uv+xy & v^2+y^2 \\
\end{bmatrix}
\neq
\begin{bmatrix}
    u^2+v^2 & ux+vy \\
    ux+vy & x^2+y^2 \\
\end{bmatrix}
$$
Therefore, $A^{T}A \neq AA^{T}$ in general.  

**Demonstration with 2 x 2 numeric valued matrices**

```{r}
# Let A be a 2 x 2 matrix
A = matrix(c(2,4,6,8), 2, 2)
A_T = t(A) # A transpose
A_TA = A_T  %*% A # A transpose multiplied by A
AA_T = A %*% A_T # A multiplied by A transpose
```

```{r}
A_TA
```

```{r}
AA_T
```

```{r}
# check if the two matrices are equal
AA_T == A_TA
```
```{r}
# check if the two matrices are equal
identical(A_TA, AA_T)
```
**Demonstration with 3 x 3 numeric valued matrices**
```{r}
A = matrix(c(1,2,3,4,5,6,7,8,9), 3, 3)
A_T = t(A)
A_TA = A_T  %*% A
AA_T = A %*% A_T
```
```{r}
A_TA
```
```{r}
AA_T
```
```{r}
# check if the two matrices are equal
AA_T == A_TA
```
```{r}
# check if the two matrices are identical
identical(A_TA, AA_T)
```

### 1.2
For a special type of square matrix A, we get $A^{T}A = AA^{T}$.  Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).  
</b>  
**Solution**  

The condition for which we get the above is for a symmetric matrix.  
```{r}
# Let A be a symmetric matrix
A = matrix(c(3,-2,4,-2,6,2,4,2,3), 3, 3)
A
```
```{r}
# check if the matrix is symmetric
isSymmetric.matrix(A)
```

```{r}
A_T = t(A)
A_TA = A_T  %*% A
AA_T = A %*% A_T
```

```{r}
A_TA
```

```{r}
AA_T
```


```{r}
# check if the two matrices are equal
AA_T == A_TA
```
```{r}
# check if the two matrices are identical
identical(A_TA, AA_T)
```
<center>
<h3> 2: Problem Set 2 </h3>
</center>
Write an R function to factorize a square matrix A into LU or LDU.
</b>  
**Solution**  
```{r}
LU_factorizer <- function(X) {
  m <- dim(X)[1]
  n <- dim(X)[2]
  L <- diag(m)
  U <- X
  
  # Check to ensure that M is a square matrix
  if (m != n) {
    return(cat("This is a not a square matrix. \nThis function works for only square matrices"))
  }

  if (m > 3 | n > 3){
    return("Please enter a square matrix of lower dimension")
  }
  
  # If dimension is 1, then U=A and L=[1]
  if (n==1 && m==1) {
    return(list(L,U))
  }
  
  # Check if zero is in diagonal of the provided matrix
  if (0 %in% diag(X)){
    return("This fuction does not work if zero is in the diagonal of the matrix")
    
  }
  
  # Loop through the lower triangle to determine the multiplier for each position:
  for(i in 2:m) {
    for(j in 1:(i-1)) {
      multiplier <- -U[i,j] / U[j,j]
      U[i, ] <- multiplier * U[j, ] + U[i, ]
      L[i,j] <- -multiplier
    }
  }
  
  return(list(L,U))
}
```


**Demonstration with a non-squared matrix**
```{r}
# Test for a non-square matrix
A_nonsquare = matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3)
LU_factorizer(A_nonsquare)
```

**Demonstration with 3 x 3 numeric valued matrices**
```{r}
# Let's declare a random 5 x 5 matrix, Y
Y = matrix(1:9,ncol=3)
Y
```
Apply the LU_factorizer function on Y
```{r}
LU = LU_factorizer(Y)
L = LU[[1]]
U = LU[[2]]
```

Display the L and U decomposition
```{r}
# display L
L
```

```{r}
U
```
```{r}
LU_result = L %*% U
LU_result
```
```{r}
# check if Y = LU
Y == LU_result
```

**Demonstration with a random 3x3 matrix**
```{r}
set.seed(94)
X <- matrix(sample(c(1:10), size = 9),nrow = 3)
X
```
Apply the LU_factorizer function on Y
```{r}
LU = LU_factorizer(X)
L = LU[[1]]
U = LU[[2]]
```

Display the L and U decomposition
```{r}
# display L
L
```

```{r}
U
```

```{r}
LU_result = L %*% U
LU_result
```
```{r}
# check if X = LU
X == LU_result
```





























