---
title: "CUNY SPS HW 7"
author: "Chinedu Onyeka"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center> **Problem 1**</center> 
Let X1, X2, . . . , Xn be n mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to k. Let Y denote the
minimum of the Xi’s. Find the distribution of Y .  
</br>
**Solution 1**  
Assuming that each $X_{i}$ has k possibilities, we will have $k^{n}$ possible values in total. 
1 is the minimum of the Xi’s, the number of ways of getting Y.  
for $P(x = 1) = \frac{k^{n} - (k-1)^{n}}{k^{n}}$  
for $P(x = 2) = \frac{(k-2+1)^{n} - (k-2)^{n}}{k^{n}}$  
For $x_{1}, x_{2}, . . . , x_{n}$ mutually independent random variables, If Y = j  then, the minimum value j is given as:  
$P(x = j) = \frac{(k-j+1)^{n} - (k-j)^{n}}{k^{n}}$  


<center> **Problem 2**</center>   
Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).
This machine has a manufacturer’s expected lifetime of 10 years. This means
that we expect one failure every ten years. (Include the probability statements
and R Code for each part.).  
</br>
*Problem 2a (Geometric)*  
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)  
**Solution 2a (Geometric)**  
*Probability of failure in 8 years*  
Given that p is the probability of success and q = 1 - p is the probability of failure for one trial, the geometric probability distribution is given by: 
$P(X=x) = p(1−p)^{x-1}$, x = 1, 2, 3, ... where x is the number of trials up to the first success

```{r}
x = 8
p = 1 / 10 # probability of success
prob_geometric = p * ((1 - p)^(8 - 1))
prob_fail_8years_geometric =  round(prob_geometric, 4)
prob_fail_8years_geometric
```
*Expected value*  
The expected value of a geometric distribution is given by $E(X) = 1/p$
```{r}
expected_value_geometric = 1/p
expected_value_geometric
```
*Standard Deviation*  
The standard deviation of a geometric distribution is given by $S.D = \frac{\sqrt{1-p}}{p}$
```{r}
sd_geom = sqrt(1 - p)/p 
sd_geometric = round(sd_geom, 4)
sd_geometric
```
</br>
*Problem 2b (Exponential)*  
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential.  
**Solution 2b (Exponential)**  
*Probability of failure in 8 years*  
Given the rate parameter $\lambda$, The exponential distribution is given by:  
$P(X=x) = \lambda e^{-\lambda x}$  
```{r}
rate = 1/10
x = 8
prob_exponetial = rate * exp(-rate * x)
prob_fail_8years_exponential = round(prob_exponetial, 4)
prob_fail_8years_exponential
```
*Expected value*  
The expected value of an exponential distribution is given by $E(X) = 1/\lambda$  
```{r}
expected_value_exponential = 1/rate
expected_value_exponential
```
*Standard Deviation*  
The standard deviation of an exponential distribution is given by $\sigma = \frac{1}{\lambda}$  
```{r}
sd_expo = 1/rate
sd_exponential = round(sd_expo, 4)
sd_exponential
```
</br>
*Problem 2c (Binomial)*  
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years)  
**Solution 2c (Binomial)**  
*Probability of 0 success in 8 years*  
Given the probability of success p, the probability of x successes in n trials is given by:  
$P(x,n,p)=\binom{n}{x}p^{x}(1−p)^{n-x}$  
```{r}
p = 1/10
prob_0_success = choose(8, 0) * p^0 * (1-p)^8
prob_0_success_in_8_years = round(prob_0_success, 4)
prob_0_success_in_8_years
```
OR
```{r}
prob = dbinom(0, 8, p)
prob_0_success_in_8_trials = round(prob, 4)
prob_0_success_in_8_trials
```
*Expected value*  
The expected value of a binomial distribution is given by $E(X) = np$  
```{r}
n = 8
p = 1/10
expected_value_binomial = n*p
expected_value_binomial
```
*Standard Deviation*  
The standard deviation of a binomial distribution is given by $\sigma = \sqrt{npq}$  
```{r}
sd_binom = sqrt(n*p*(1-p))
sd_binomial = round(sd_binom, 4)
sd_binomial
```
</br>
*Problem 2d (Poison)*  
What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson.  
**Solution 2d (Poison)**  
*Probability of failure in 8 years*  
The poison distribution formula is given by:  
$P(X=x) = \frac{\lambda ^{x}}{x!}e^{-\lambda}$  
$\lambda = 10$  

```{r}
rate = 8 * (1 / 10)
e <- exp(1)
prob_poison = (rate^0 * exp(-rate)) / factorial(0)
prob_fail_8years_poison = round(prob_poison, 4)
prob_fail_8years_poison
```
OR
```{r}
prob_pois = dpois(0, rate)
prob_poison_8years = round(prob_pois, 4)
prob_poison_8years
```
*Expected value*  
The expected value of the poison distribution is given by: $E(X) = \lambda$
```{r}
expected_value_poison <- rate
expected_value_poison
```
*Standard Deviation*  
The standard deviation of the poison distribution is given by:  
$\sigma = \sqrt{\lambda}$
```{r}
sd_pois = sqrt(rate)
sd_poison = round(sd_pois, 4)
sd_poison
```
















