---
title: "CUNY SPS DATA 605 HW 15"
author: "Chinedu Onyeka"
date: "2023-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libs, warning=FALSE, message=FALSE}
library(tidyverse)
```



<center><h3>Problem 1</h3></center>
Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.  
$(5.6, 8.8),\ (6.3, 12.4),\ (7, 14.8),\ (7.7, 18.2),\ (8.4, 20.8)$  

<h4> Solution 1 </h4>

```{r}
x = c(5.6, 6.3, 7, 7.7, 8.4)
y = c(8.8, 12.4, 14.8, 18.2, 20.8)
regression_model = lm(y ~ x)
regression_model
```
From the results of the regression model, the equation of the regression line for the given points is:
$y = -14.8 + 4.26x$  
<br>

<center><h3>Problem 2</h3></center>
Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form
$(x, y, z)$. Separate multiple points with a comma.  
$f(x, y) = 24x - 6xy^2 - 8y^3$

<h4> Solution 2 </h4>
Find the partial derivatives of $f(x,y):$  

$\frac{\partial}{\partial x}f(x, y) = 24 - 6y^2$  

$\frac{\partial}{\partial y}f(x, y) = -12xy - 24y^2$  

Let, $24-6y^2=0; \ then,\ y^2 = 4 \ => y = \pm \ 2$  

For $y=2, \ then\ -12xy - 24y^2=0; \ -24x = 24 \times 4; \ => x=-4$  


For $y=-2, then\ -12xy - 24y^2=0; \ 24x = 24\times 4; \ => x=4$.

Calculating $f(x, y)$ using these values of x and y  

$f(4,-2) = 24\times 4 - 6\times4\times (-2)^2 - 8 \times (-2)^3 = 64$  

$f(-4,2) = 24\times (-4) - 6\times(-4)\times 2^2 - 8 \times 2^3 = -64$  

Therefore, the two critical points are: $(4,-2,64)\  and\  (-4, 2, -64)$ which are two points in a 3-dimensional space/surface.  

Find the second partial derivatives and use that to test if points are minimum, maximum or saddle points.  

Second partial derivatives:

$\frac{\partial^2}{\partial x}f(x, y) = 0$  

$\frac{\partial^2}{\partial y}f(x, y) = -12x-48y$  

$\frac{\partial^2}{\partial x \partial y}f(x, y) = -12y $. This is the mmixed partial derivative.  

Then find $D(x,y) = f_{xx} f_{yy} - f^2_{xy} = 0 - (-12y)^2 = -144y^2$. This will always result in D < 0 for any value of y.  

According to the second derivative test, any critical point will be a saddle point since $D(x, y)<0\  \forall \ (x, y)\  \epsilon \ R $. 
<br>  

<center><h3>Problem 3</h3></center>
A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for $x$ dollars and the "name" brand for $y$ dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.  

Step 1. Find the revenue function $R ( x, y )$.  
Step 2. What is the revenue if she sells the "house" brand for \$2.30 and the "name" brand for \$4.10?  


<h4> Solution 3 </h4>  
*Step 1*  
Combine the two functions by multiplying the price with the number of units.

$$
\begin{split}
R(x,y) &= (81 - 21x + 17y)x + (40 + 11x - 23y)y \\
&= 81x - 21x^2 + 17xy + 40y + 11xy - 23y^2\\
&=81x + 40y + 28xy - 21x^2 - 23y^2
\end{split}
$$
Therefore, the revenue function is given by: $R ( x, y ) = 81x + 40y + 28xy - 21x^2 - 23y^2$.  

*Step 2*  
Substitute, $x = 2.3, \ and\  y = 4.10$  
$R(2.3, 4.1) = 81\times 2.3 + 40\times 4.1 + 28\times 2.3\times 4.1 - 21\times (2.3)^2 - 23\times (4.1)^2 = 116.62$
<br>  

<center><h3>Problem 4</h3></center>
A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a
product each week. The total weekly cost is given by $C(x, y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700$, where $x$ is the number of units produced in Los Angeles and $y$ is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?


<h4> Solution 4 </h4>  
From the given problem, we can see that sum of units in Los Angeles and Denver is 96 units.  
Therefore, $x + y = 96; => x = 96 - y$. Substituting this into the cost function, we have:  
$$
\begin{split}
C(x,y) = C(96-y,y) &= \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700 \\
& = \frac{1}{6} (96-y)^2 + \frac{1}{6} y^2 + 7\times (96-y) + 25y + 700 \\
& = \frac{1}{6}(y^2 - 192 y + 9216) + \frac{1}{6}y^2 + 672 - 7y + 25y + 700\\
& = \frac{1}{6}y^2 - 32y + 1536 + \frac{1}{6}y^2 + 18y + 1372\\
& = \frac{1}{3}y^2 - 14y + 2908\\
& = f(y)
\end{split}
$$

Now we have the cost as a function of y only. To minimize the cost, we find the first derivative of the cost function wrt y and set the derivative to zero.  
$f'(y) = \frac{2}{3}y-14; \ y = 21$. Therefore, $x = 96-y = 75$.  
Hence, the company should manufacture 75 units in Los Angeles and 21 units in Denver to minimize cost.  
<br>

<center><h3>Problem 5</h3></center>

Evaluate the double integral on the given region.

$$
\int\int_R (e^{8x+3y}) dA,\ \  R:2\le x\le4\ and\ 2 \le y \le 4
$$  
Write your answer in the exact form without decimals.  

<h4> Solution 5 </h4> 
For the double integral, we integrate first with respect to y and apply the integral limits for y and then do the same for x as well as shown below:  

$$
\begin{split}
\int_2^4\int_2^4 (e^{8x+3y})\ dy\ dx &= \int_2^4 (\frac{1}{3}e^{8x+3y})|_2^4\ dx\\
&= \int_2^4 ((\frac{1}{3}e^{8x+12})-(\frac{1}{3}e^{8x+6}))\ dx\\
&= \int_2^4 \frac{1}{3}e^{8x+6}(e^6-1)\ dx\\
&=  \frac{1}{24}e^{8x+6}(e^6-1) |_2^4\\
&= \frac{1}{24}e^{32+6}(e^6-1)-\frac{1}{24}e^{16+6}(e^6-1)\\
&= \frac{1}{24}(e^6-1)(e^{38}-e^{22})\\
&= \frac{1}{24}(e^{44} - e^{38} - e^{28} + e^{22})
\end{split}
$$