---
title: "CUNY SPS HW11"
author: "Chinedu Onyeka"
date: "2023-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<center><h3>Question</h3></center>
Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and
replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and
residual analysis.)

<center><h3>Solution</h3></center>

*Load required libraries*
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
```

<b>Take a glimpse at the dataset</b>

```{r}
glimpse(cars)
```
We can see that the dataset has 50 rows and 2 columns (speed and dist).  

<b>Summary for the dataset</b>
```{r}
summary(cars)
```
<b> Visualize the data </b>

```{r warning=FALSE}
p = ggplot(cars, aes(x=speed, y=dist)) + geom_point() + theme(panel.grid.major = element_line(colour = "lemonchiffon3"),
    panel.grid.minor = element_line(colour = "lemonchiffon3"),
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(family = "sans",
        size = 11), axis.text.y = element_text(family = "sans",
        size = 11), plot.title = element_text(size = 15,
        hjust = 0.5), panel.background = element_rect(fill = "gray85"),
    plot.background = element_rect(fill = "antiquewhite")) +labs(title = "Stopping Distance vs Speed",
    x = "Speed", y = "Stopping Distance")
p
```

From the scatter plot, as the speed increases, the stopping distance tend to increase as well.  

<b> Build the Model </b>
```{r}
lm_cars = lm(cars$dist ~ cars$speed)
lm_cars
```
From the output of the model, the linear function is given by:  
$dist = 3.932 * speed - 17.579$  

<b>Evaluate the Model - Quality Evaluation</b>  
```{r}
# Check the summary of the model
summary(lm_cars)
```
The values from the summary show that there is little variability in the data and the result is statistically significant.  

<b>Residual Analysis</b>  
*Residual vs Fitted plot*
```{r}
plot(fitted(lm_cars),resid(lm_cars), main="Residuals vs Fitted", xlab = "Fitted", ylab = "Residuals")
abline(0, 0)
```

From this plot, we can say that there seems to be constant variance for the residuals and the data is not heteroscedacity.  

*QQ Plot*
```{r}
qqnorm(resid(lm_cars))
qqline(resid(lm_cars))
```

From the Q-Q plot, we can see that the residuals follow a nearly normal distribution.  

*Entire plot*
```{r}
par(mfrow=c(2,2))
plot(lm_cars)
```

From the residual analysis, we can see that the conditions for linear regression are satisfied for this data and it can be fitted by using a linear model approach. Although I am not satisfied with the R-squared value of about 0.65, acceptable values for the R-squared value is highly dependent on the use case of the model.












