---
title: "CUNY SPS Discussion 11"
author: "Chinedu Onyeka"
date: "2023-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<center><h3>Question</h3></center>
Using the [salary_dataset from kaggle](https://www.kaggle.com/datasets/abhishek14398/salary-dataset-simple-linear-regression), build a linear model for salary vs years of experience. Conduct residual analysis. Was the linear model appropriate? Why or why not? 

<center><h3>Solution</h3></center>

*Load required libraries*
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
```

<b>Read the data and take a glimpse at the dataset</b>  

*Read the data*
```{r warning=FALSE, message=FALSE}
salary_url = "https://raw.githubusercontent.com/chinedu2301/data605-computer-maths/main/discussions/discussion11/salary_dataset.csv"
# read the data from github url
salary_dataset_raw = read_csv(salary_url) 
```

*Take a glimpse at the data*
```{r}
salary_dataset = salary_dataset_raw %>% select("YearsExperience", "Salary")
glimpse(salary_dataset)
```

We can see that the dataset has 30 rows and 2 columns (Salary and YearsExperience).  

<b>Summary for the dataset</b>
```{r}
summary(salary_dataset)
```
<b> Visualize the data </b>

```{r warning=FALSE}
p = ggplot(salary_dataset, aes(x=YearsExperience, y=Salary)) + geom_point() + 
    theme(panel.grid.major = element_line(colour = "lemonchiffon3"),
    panel.grid.minor = element_line(colour = "lemonchiffon3"),
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(family = "sans",
        size = 11), axis.text.y = element_text(family = "sans",
        size = 11), plot.title = element_text(size = 15,
        hjust = 0.5), panel.background = element_rect(fill = "gray85"),
    plot.background = element_rect(fill = "antiquewhite")) +labs(title = "Salary vs YearsExperience",
    x = "YearsExperience", y = "Salary")
p
```

From the scatter plot, as the YearsExperience increases, the Salary tend to increase as well. There is a clear linear relationship between the Salary and YearsExperience in the given data. 

<b> Build the Model </b>
```{r}
lm_salary = lm(salary_dataset$Salary ~ salary_dataset$YearsExperience)
lm_salary
```
From the output of the model, the linear function is given by:  
$Salary = 9450 * YearsExperience - 24848$  

<b>Evaluate the Model - Quality Evaluation</b>  
```{r}
# Check the summary of the model
summary(lm_salary)
```
The values from the summary show that there is little variability in the data and the result is statistically significant.  

<b>Residual Analysis</b>  
*Residual vs Fitted plot*
```{r}
plot(fitted(lm_salary),resid(lm_salary), main="Residuals vs Fitted", xlab = "Fitted", ylab = "Residuals")
abline(0, 0)
```

From this plot, we can say that there seems to be constant variance for the residuals and the data is not heteroscedastic.  

*QQ Plot*
```{r}
qqnorm(resid(lm_salary))
qqline(resid(lm_salary))
```

From the Q-Q plot, we can see that the residuals follow a nearly normal distribution.  

*Entire plot*
```{r}
par(mfrow=c(2,2))
plot(lm_salary)
```

From the residual analysis, we can see that the conditions for linear regression are satisfied for this data and it can be fitted by using a linear model approach. The R-squared value of about 0.957, acceptable values for the R-squared value is highly dependent on the use case of the model. From the results of the model, I believe that it is appropriate to use a simple linear regression model for this data












