---
title: "CUNY SPS HW 12"
author: "Chinedu Onyeka"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center><h3>Question</h3></center>
The attached who.csv dataset contains real-world data from 2008. The variables included follow:  
- `Country`: name of the country
- `LifeExp`: average life expectancy for the country in years
- `InfantSurvival`: proportion of those surviving to one year or more
- `Under5Survival`: proportion of those surviving to five years or more
- `TBFree`: proportion of the population without TB
- `PropMD`: proportion of the population who are MDs
- `PropRN`: proportion of the population who are RNs
- `PersExp`: mean personal expenditures on healthcare in US dollars at average exchange rate
- `GovtExp`: mean government expenditures per capita on healthcare, US dollars at average exchange rate
- `TotExp`: sum of personal and government expenditures  


<center><h3>Solution</h3></center>

*Load required libraries*
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
```

<b>Read the Data and take a glimpse at the dataset</b>

```{r read-data, message=FALSE, warning=FALSE}
url_who <- ""
who_raw <- read_csv(url_who)
```


```{r}
glimpse(cars)
```

<center><h3>Question 1</h3></center>
Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the
variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss
whether the assumptions of simple linear regression met.
