---
title: "CUNY SPS Discussion 12"
author: "Chinedu Onyeka"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center><h3>Question</h3></center>
Using R, build a multiple regression model for data that interests you.  Include in this model at least one quadratic term, one dichotomous term, and one dichotomous vs. quantitative interaction term.  Interpret all coefficients. Conduct residual analysis.  Was the linear model appropriate? Why or why not? 
<br>

<center><h3>Solution</h3></center>

*Load required libraries*
```{r library, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
```

<b>Read the Data into memory</b>

```{r read-data, message=FALSE, warning=FALSE}
data(trees)
```

<b>Display the data</b>
The data is an in-built data provided by R. It is about Diameter, Height and Volume for Black Cherry Trees.  

```{r display-data}
trees_display = kable(trees) %>%
                  kable_paper("hover", full_width = F) %>%
                  scroll_box(width = "850px", height = "350px")
trees_display
```

<br>
<b>Glipmse of the data</b>
```{r glimpse}
glimpse(trees)
```
<br>
<b>Summary of the data</b>
```{r summary}
summary(trees)
```

```{r transform-data}
trees_transformed <- trees %>%
                  mutate(Height2 = Height ^ 2,  # quadratic term
                  HoursCategory = ifelse(Height > mean(Height), 1.0, 0.0),  #dichotomous term
                  Height_Girth = Height * Girth  #quantitative interaction term
  )
```

 
<b> Scatter Plot - Q4</b>
```{r scatter-plot}
p = ggplot(trees_transformed, aes(x=(Height2 + HoursCategory + Height_Girth), y = Volume)) + geom_point() + theme_minimal() +
    theme(panel.grid.major = element_line(colour = "lemonchiffon3"),
    panel.grid.minor = element_line(colour = "lemonchiffon3"),
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(family = "sans",
        size = 11), axis.text.y = element_text(family = "sans",
        size = 11), plot.title = element_text(size = 15,
        hjust = 0.5), panel.background = element_rect(fill = "gray85"),
    plot.background = element_rect(fill = "antiquewhite")) +labs(title = "Volume - Multi Regression",
    x = "Height2 + HoursCategory + Height_Girth", y = "Volume")
p
```

<br>
<b> Simple Linear Regression - Q4 </b>  
```{r 4-simple-linear-regression}
lm_tree <- lm(Volume ~ Height2 + HoursCategory + Height_Girth, data = trees_transformed)
summary(lm_tree)
```

<b> Residual vs Fitted - Q4</b>
```{r 4-residual-fitted}
plot(fitted(lm_tree),resid(lm_tree), main="Residuals vs Fitted", xlab = "Fitted", ylab = "Residuals")
abline(0, 0)
```

<br>
<b>Q-Q Plot</b>
```{r 4-qq-plot}
qqnorm(resid(lm_tree))
qqline(resid(lm_tree))
```
<br>  
<b> Interpret the F statistics, R^2, standard error,and p-values only:</b>  
F statistic: 270.5 on 3 and 27 DF  
p-value: < 2.2e-16 - The p-value being less than 0.05 means that the result is statistically significant.
R-squared:  0.9678 - This means that the model accounts for about 97% of variability in the data.  
Residual standard error: The TotExp can deviate from the regression line by approximately 3.109 on 27 degrees of freedom 
<br>
There are four (4) main assumptions for Linear Regression and they are:  
<li>Linearity: The relationship between X and Y must be linear. As can be seen from the scatter plot above, `Height2 + HoursCategory + Height_Girth)` apear to have a linear relationship and this condition is satisfied.</li>  
<li>Homoscedacity: There should be constant variance in the residuals. From the Residual vs Fitted Plot shown above, it seem to appear that there is a constant variance and thus the homoscedacity criterion is satisfied. </li>  
<li>Normality: The data should be normally distributed. From the Q-Q plot shown above, the data is nearly normal.</li>  
<li>Independence: The observations should be independent of each other. This may be difficult to determine from looking at the data and we may have to rely on the assumptions provided by the data collector. From the residual plot, we can assume that this criteria is met. 
Since the Linearity, Homoscedacity, and Normality conditions are satisfied, we can conclude that the assumptions for Linear Regression are met.</li>  
<br>

Even though this model might suffer from multi-collinearity, in general, a linear model seem to be appropriated for the data.













