---
title: "DATA 605 Homework3"
author: "Chinedu Onyeka"
date: "2023-09-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Questions: ** 
<center>
<h3> 1: Problem Set 1 </h3>
</center>
#### 1.1 What is the rank of the matrix A?

$$
A = \begin{bmatrix}
1 & 2 & 3 & 4 \\
-2 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3 \\
\end{bmatrix}
$$  
</b>  
**Solution**  

```{r}
A <- matrix(c(1,-2,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3), ncol=4)
A
```
```{r}
# rank of the matrix, A
rankA <- qr(A)$rank
rankA
```
Therefore the rank of the matrix is 4.  

#### 1.2 Given an $m \times n$ matrix where $m>n$, what can be the maximum rank?
The minimum rank, assuming that the matrix is non-zero?  
</br>
**Solution**  
For a non-zero matrix, the minimum and maximum rank is given by:  

Minimum rank =  1  
Maximum rank = n  

#### 1.3 What is the rank of matrix $B$?

$$
B = 
\begin{bmatrix}
    1 & 2 & 1 \\
    3 & 6 & 3 \\
    2 & 4 & 2 \\
\end{bmatrix}
$$ 
  
</br>
**Solution**
Looking at the matrix, the rank is 1 given than there is only one linearly independent row for the matrix, B. The rank of a matrix is simply the number of linearly independent rows of the matrix. Given that matrix B has rows 2 and 3 as multiples of row 1, it means that there is only one linearly independent row for the matrix and thus the rank is 1.

```{r}
# checking the matrix using R code:
B <- matrix(c(1,3,2,2,6,4,1,3,2), nrow = 3)
B
```
```{r}
# rank of matrix B
rankB <- qr(B)$rank
rankB
```

<center>
<h3> 2: Problem Set 2 </h3>
</center>

Compute the eigenvalues and eigenvectors of the matrix $A$. You'll need to show your work. You'll need to write out the characteristic
polynomial and show your solution.

$$
A = 
\begin{bmatrix}
    1 & 2 & 3 \\
    0 & 4 & 5 \\
    0 & 0 & 6 \\
\end{bmatrix}
$$

</br>

**Solution**  
The eigenvalue(s) is $\lambda$ is defined as $Ax = \lambda x$ where $x$ is the eigen vector associated with the eigenvalue $\lambda$.  
$|A-\lambda I| = 0$ where $I$ is a $3 \times 3$ identity square matrix.

Therefore:
$$
\begin{bmatrix}
    1 & 2 & 3 \\
    0 & 4 & 5 \\
    0 & 0 & 6 \\
\end{bmatrix}
- \lambda
\begin{bmatrix}
    1 & 0 & 0 \\
    0 & 1 & 0 \\
    0 & 0 & 1 \\
\end{bmatrix} = 
\begin{bmatrix}
    1-\lambda & 2 & 3 \\
    0 & 4-\lambda & 5 \\
    0 & 0 & 6-\lambda \\
\end{bmatrix}
$$

Find the determinant of the resulting matrix and set it equal to zero to find the values of $\lambda$  

$$
det
\begin{bmatrix}
    1-\lambda & 2 & 3 \\
    0 & 4-\lambda & 5 \\
    0 & 0 & 6-\lambda \\
\end{bmatrix} = 0
$$  

Multiplying along the first column to take advantage of the zeros in the first columns:  
$(1-\lambda)[(4-\lambda)(6-\lambda) - 0] = 0$  
$=> (1-\lambda)(4-\lambda)(6-\lambda) = 0$ This is the characteristic polynomial and it's better to keep it unexpanded since it makes it a lot easier to see the values of $\lambda$.  

Therefore $\lambda = 1, 4, 6$.  
Thus, the eigenvalues are $1, 4,$ and $6$.  

```{r}
# verify using R
A <- matrix(c(1, 0, 0, 2, 4, 0, 3, 5, 6), nrow=3, ncol=3)
eigen_values <- eigen(A)
eigen_values$values
```  



#### **Eigen-vectors**.  
</br>
**Eigen-vector for $\lambda = 1$**;  
$(A - \lambda I)x = 0$; $x$ is the eigen-vector   

For $\lambda = 1$;  
$$
\begin{bmatrix}
    1-1 & 2 & 3 \\
    0 & 4-1 & 5 \\
    0 & 0 & 6-1 \\
\end{bmatrix} 
\times
\begin{bmatrix}
    u \\
    v \\
    w \\
\end{bmatrix}  = 0
$$

$$
\begin{bmatrix}
    0 & 2 & 3 \\
    0 & 3 & 5 \\
    0 & 0 & 5 \\
\end{bmatrix} 
\times
\begin{bmatrix}
    u \\
    v \\
    w \\
\end{bmatrix}  = 0; 
$$
 $5w = 0; w = 0$; $3v + 5w = 0; 3v + 5(0) = 0; v = 0$; $0u + 2v + 3w = 0; 0 + 0 + 0 = 0; 0=0$  
 This gives a trivial solution for x;  
 
Thus the eigen-vector associated with $\lambda = 1$ is given by:    
$$ 
 x = u
\begin{bmatrix}
    1 \\
    0 \\
    0 \\
\end{bmatrix} 
$$

$\forall$ $u$ $\epsilon$ $\mathbb{C}$  
</br>
**Eigen-vector For $\lambda = 4$**  
$$
\begin{bmatrix}
    1-4 & 2 & 3 \\
    0 & 4-4 & 5 \\
    0 & 0 & 6-4 \\
\end{bmatrix} 
\times
\begin{bmatrix}
    u \\
    v \\
    w \\
\end{bmatrix}  = 0
$$
$$
\begin{bmatrix}
    -3 & 2 & 3 \\
    0 & 0 & 5 \\
    0 & 0 & 2 \\
\end{bmatrix} 
\times
\begin{bmatrix}
    u \\
    v \\
    w \\
\end{bmatrix}  = 0; 
$$  

$2w = 0; w=0$; $0v + 2w = 0; 0v + 0 + 0 = 0; 0 = 0$ => $v$ appears to result in a trivial solution.  
$-3u + 2v + 3w = 0; -3u + 2v + 0 = 0$; $u = \frac{2}{3}v$.  
Thus the eigen-vector associated with $\lambda = 4$ is given by:  

$$ 
 x = v
\begin{bmatrix}
    \frac{2}{3} \\
    1 \\
    0 \\
\end{bmatrix} 
$$  

$\forall$ $v$ $\epsilon$ $\mathbb{C}$  
</br>
**Eigen-vector For $\lambda = 6$**  
$$
\begin{bmatrix}
    1-6 & 2 & 3 \\
    0 & 4-6 & 5 \\
    0 & 0 & 6-6 \\
\end{bmatrix} 
\times
\begin{bmatrix}
    u \\
    v \\
    w \\
\end{bmatrix}  = 0
$$  
$$
\begin{bmatrix}
    -5 & 2 & 3 \\
    0 & -2 & 5 \\
    0 & 0 & 0 \\
\end{bmatrix} 
\times
\begin{bmatrix}
    u \\
    v \\
    w \\
\end{bmatrix}  = 0; 
$$ 

Clearly, the last row provides a trivial solution.  
$-2v + 5w = 0; v = \frac{5}{2}w; -5u + 2v + 3w = 0$; $-5u + 2\times\frac{5}{2}w + 3w = 0$; $-5u + 5w + 3w = 0$; $u = \frac{8}{5}w$  
Thus the eigen-vector associated with $\lambda = 6$ is given by:  

$$ 
 x = w
\begin{bmatrix}
    \frac{8}{5} \\
    \frac{5}{2} \\
    1 \\
\end{bmatrix}
$$  

$\forall$ $w$ $\epsilon$ $\mathbb{C}$  

Hence, the eigen vectors corresponding to $\lambda = 1, 4, 6$ are:  


$$ 
 x = u
\begin{bmatrix}
    1 \\
    0 \\
    0 \\
\end{bmatrix}; 
 v
\begin{bmatrix}
    \frac{2}{3} \\
    1 \\
    0 \\
\end{bmatrix};  
w
\begin{bmatrix}
    \frac{8}{5} \\
    \frac{5}{2} \\
    1 \\
\end{bmatrix}
$$
$\forall$ $u, v, w$ $\epsilon$ $\mathbb{C}$ respectively.













